Formulating and Validating RDF Constraints Generically

### 2014-12-31 paper/paper 2/paper 2.tex
constraint classification

The \textbf{contributions} of this paper are:
\begin{itemize}
	%\item Any specific constraint (expressed by any constraint language) can be transformed into a generic constraint (expressed by the generic constraint language)
	%\item Any specific constraint (expressed by any constraint language A) can be transformed into any specific constraint (expressed by any constraint language B)
	%\item Any specific constraint (expressed by any constraint language) can be validated automatically (if an automatic validation is defined only once for corresponding generic constraint)
	\item We define a terminology for RDF constraints and constraining languages
	\item We developed an ontology describing any RDF constraint which can also be expressed by any constraint language
	\item We present that any RDF constraint can be described by the developed ontology - whether the constraint can be expressed in DL or not 
	\item We show how to transform specific constraints (expressed by any constraint language A) into generic constraints and into specific constraints (expressed by any other constraint language B)
  \item We explain that any specific constraint can be validated automatically (if an automatic validation is defined only once for the corresponding generic constraint)
\end{itemize}

@@

\section{An Ontology Describing RDF Constraints Generically} 
\label{sec:ontology}

We developed a very simple ontology (three classes, three object properties, and three data properties) which is universal enough to describe any RDF constraint expressible by any RDF constraint language.
Therefore, we do not invent a new constraint language.
We distinguish between two sub-types of \ms{constraint}: 
\begin{itemize}
	\item \ms{property constraints} and 
	\item \ms{class constraints}. 
\end{itemize}

For both property and class constraints a context class, a list of classes, the constraining element, and the constraining value can be stated
(see figure \ref{fig:generic-constraint-language-ontology}). Lists of left and right properties can only be stated for property constraints.

\begin{figure}
	\centering
		\includegraphics[width=0.80\textwidth]{images/generic-constraint-language-ontology.png}
	\caption{generic constraint language ontology}
	\label{fig:generic-constraint-language-ontology}
\end{figure}

In order to define this ontology, it is needed to define the terminology for the formulation of constraints. 
A \ms{constraint language} is a language which is used to formulate constraints.
The W3C Data Shapes WG defines \ms{constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
A constraint holds for all individuals of a context class (\ms{contextClass}).
Consider, e.g., the following cardinality restriction on the property \ms{commandsVessel}, which restricts that a captain must command at least one vessel: 
\begin{DL}
$Captain \sqsubseteq \geq1 commandsVessel . Vessel $
\end{DL}
Mapped to the GCLO the cardinality restriction is classified as a property constraint which holds for individuals of the class \ms{Captain} - the context class:
\begin{gcotable}
property & Captain & commandsVessel & - & Vessel & $\geq$ & 1 \\
\end{gcotable}
The constraining element (\ms{constrainingElement}) is the operator indicating the actual type of constraint like DL concept and role constructors, (in)equality, and further keywords for constrains which cannot be expressed in DL (e.g. regular expressions, constraining facets).
In some cases, a constraint is only complete when in addition to the constraining element a constraining value (\ms{constrainingValue}) is stated.
The cardinality restriction 
\begin{DL}
$\geq$1 commandsVessel . Vessel
\end{DL}
constructs an anonymous class of all individuals who command at least one vessel.
Thus, the constraining element of this property constraint is the DL at-least restriction \ms{$\geq$} and the constraining value is \ms{1}.
\ms{Classes} indicate the list of classes constraints refer to.
In case of the property constraint above, the qualified cardinality restriction refers to the class \ms{Vessel}, 
i.e., restricting the objects, the property \ms{commandsVessel} point to, to be of the class \ms{Vessel}.

\section{Complex Constraints}

\ms{Complex constraints} encompass complex constraints and/or \ms{simple constraints} (atomic constraints).
DL statements of complex constraints are created out of DL statements of composed constraints.

\textbf{Context-Specific Exclusive OR of Property Groups}\footnote{corresponds to R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC}\textbf{.}
Exclusive or is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
Only one of multiple property groups should be present in the data.
Half-Klingons, e.g., either have a klingon mother and a human father or a human mother and a klingon father, which can be expressed by ShEx:

@@

@@

\section{Syntactic Sugar}

\begin{itemize}
	\item syntactic sugar for some constraints
\end{itemize}

@@

\section{Simple Constraints Not Expressible in DL}

@@

\section{Evaluation}

>>>>

The \textbf{contributions} of this paper are:
\begin{itemize}
	%\item Any specific constraint (expressed by any constraint language) can be transformed into a generic constraint (expressed by the generic constraint language)
	%\item Any specific constraint (expressed by any constraint language A) can be transformed into any specific constraint (expressed by any constraint language B)
	%\item Any specific constraint (expressed by any constraint language) can be validated automatically (if an automatic validation is defined only once for corresponding generic constraint)
	\item We define a basic terminology and classification system for RDF constraints
	\item We developed an ontology describing any RDF constraint generically (whether expressible in DL or not)%which can also be expressed by any constraint language
	%\item We present that any RDF constraint can be described by the developed ontology - whether the constraint can be expressed in DL or not 
	\item We show how to transform specific constraints (expressed by any constraint language A) into generic constraints and into specific constraints (expressed by any other constraint language B)
  \item We explain that any specific constraint can be validated automatically (if an automatic validation is defined only once for the corresponding generic constraint)
\end{itemize}

@@

\section{RDF Constraints Ontology} 
\label{sec:ontology}

In order to define this ontology to describe RDF constraints generically, it is needed to define the terminology for the formulation of constraints. 
A \ms{constraint language} is a language which is used to formulate constraints.
The W3C Data Shapes WG defines \ms{constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
We identified four dimensions to classify constraints:
\begin{itemize}
  \item \ms{Universality:} specific constraints vs. generic constraints
	\item \ms{Complexity:} simple constraints vs. complex constraints
	\item \ms{Context:} property constraints vs. class constraints
	\item \ms{DL Expressivity:} constraints expressible in DL vs. constraints not expressible in DL
\end{itemize}

As there are already five promising constraint languages, our purpose is not to invent a new constraint language.
We developed a very simple ontology (only three classes, three object properties, and three data properties) which is universal enough to describe any RDF constraint expressible by any RDF constraint language (see the conceptual model in figure \ref{fig:RDF-CO-conceptual-model}).
We call this ontology the \ms{RDF constraints ontology (RDF-CO)}\footnote{Available at: \url{https://github.com/boschthomas/RDF-constraints-ontology}}.

\begin{figure}
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CO-conceptual-model.png}
	\caption{RDF constraints ontology (RDF-CO) - conceptual model}
	\label{fig:RDF-CO-conceptual-model}
\end{figure}

\ms{Specific constraints} are expressed by specific constraint languages like DSP, OWL 2, ReSh, ShEx, and SPIN.
\ms{Generic constraints} are expressed by the RDF-CO.
As RDF-CO describes constraints generically, RDF-CO does not distinguish constraints according to the dimension \ms{universality}. 
The majority of constraints can be expressed in DL (see section \ref{sec:evaluation}).
In contrast, there are constraints which cannot be expressed in DL, but are also representable by the RDF-CO. 
\ms{Complex constraints} encompass \ms{simple constraints} (atomic constraints) and/or further complex constraints.
DL statements representing complex constraints are created out of DL statements representing composed constraints (if expressible in DL). 
Simple constraints may be applied to either properties \ms{properties constraints} or classes \ms{class constraints}.
There are no terms representing simple and complex constraints in the RDF-CO, since context classes (associated simple constraints hold for individuals of these classes) of simple constraints may just be reused by further constraints.
As a consequence, the distinction of property and class constraints is sufficient to describe all possible RDF constraints.

\subsection{Simple Constraints (Expressible in DL)}

Sub-classes of \ms{simple constraints} are \ms{property constraints} and \ms{class constraints} (see the RDF-CO implementation model in figure \ref{fig:RDF-CO-implementation-model}). 
For both property and class constraints a context class, a list of classes, the constraining element, and the constraining value can be stated. 
Lists of left and right properties can only be stated for property constraints.

\begin{figure}
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CO-implementation-model.png}
	\caption{RDF constraints ontology (RDF-CO) - implementation model}
	\label{fig:RDF-CO-implementation-model}
\end{figure}

A constraint holds for all individuals of a context class (\ms{contextClass}).
Consider, e.g., the following cardinality restriction on the property \ms{commandsVessel}, which restricts that a captain must command at least one vessel: 
\begin{DL}
$Captain \sqsubseteq \geq1 commandsVessel . Vessel $
\end{DL}
Mapped to the GCLO the cardinality restriction is classified as a property constraint which holds for individuals of the class \ms{Captain} - the context class:
\begin{gcotable}
property & Captain & commandsVessel & - & Vessel & $\geq$ & 1 \\
\end{gcotable}
The constraining element (\ms{constrainingElement}) is the operator indicating the actual type of constraint like DL concept and role constructors, (in)equality, and further keywords for constrains which cannot be expressed in DL (e.g. regular expressions, constraining facets).
In some cases, a constraint is only complete when in addition to the constraining element a constraining value (\ms{constrainingValue}) is stated.
The cardinality restriction 
\begin{DL}
$\geq$1 commandsVessel . Vessel
\end{DL}
constructs an anonymous class of all individuals who command at least one vessel.
Thus, the constraining element of this property constraint is the DL at-least restriction \ms{$\geq$} and the constraining value is \ms{1}.
\ms{Classes} indicate the list of classes constraints refer to.
In case of the property constraint above, the qualified cardinality restriction refers to the class \ms{Vessel}, 
i.e., restricting the objects, the property \ms{commandsVessel} point to, to be of the class \ms{Vessel}.

\subsection{Simple Constraints (Not Expressible in DL)}

\subsection{Complex Constraints}

\ms{Complex constraints} encompass simple constraints and/or complex constraints.
\textbf{Context-Specific Exclusive OR of Property Groups}\footnote{corresponds to R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC}\textbf{.}
Exclusive or is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
Only one of multiple property groups should be present in the data.
Half-Klingons, e.g., either have a klingon mother and a human father or a human mother and a klingon father, which can be expressed by ShEx:

@@

\subsection{Simple Constraints (Syntactic Sugar)}

Almost 15 percent of all RDF constraints are complex constraints which can be simplified and therefore formulated as simple constraints when used in terms of syntactic sugar (see section \ref{sec:evaluation}).

@@

@@

@@

\section{Evaluation}
\label{sec:evaluation}
